---
title: "Oconee Forest Park Plant Diversity"
author: "Raymond Schneider"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE}
library(tidyverse)
library(collapsibleTree)
library(viridis)
library(kableExtra)
```

```{r load and clean plant data}
plantData <- read_csv("IAIA_Science.csv")
colnames(plantData) <- str_replace_all(colnames(plantData), " ", "_")
```

Get relative abundance and richness

```{r relative abundance and richness}
# relative abundance:
plantDataEnv <- plantData %>% group_by(Environment, Common_name) %>%
  summarize(Count = mean(Amount)) %>% # average abundance of each species across plots
  ungroup() %>%
  mutate(Count = round(Count, 0))
plantDataEnvSum <- plantDataEnv %>% group_by(Environment) %>% summarize(Total_count = sum(Count)) # get total species abundance per environment
plantDataEnvTotal <- plantDataEnv %>% left_join(plantDataEnvSum) %>%
  mutate(Rel_abundance_env = Count/Total_count) # calculate relative abundance per environment

# richness:
plantRichnessPlot <- plantData %>% group_by(Environment, Plot) %>% summarize(Plot_richness = length(unique(Common_name)))

# join everything together for plotting
plantDataTotal <- plantData %>% left_join(plantRichnessPlot) %>% left_join(plantDataEnvTotal) %>% select(!Notes) %>% unique()
```

Plot average species richness in each environment:

```{r plot spp richness, echo=FALSE}
p1 <- ggplot(plantDataTotal, aes(x = Environment, y = Plot_richness, fill = Environment)) +
  geom_boxplot(alpha=.6) +
  geom_point(show.legend = F) +
  theme_classic() +
  scale_fill_viridis_d() +
  labs(x = "Environment", y = "Species Richness")
ggsave("richness.png", p1, dpi = 300)
p1
```
Shannon-Weiner diversity index: measures diversity in terms of richness and evenness.

Write a function that calculates Shannon diversity:

```{r shannon function}
shannon <- function(x){
  P <- x/sum(x)
  index <- -sum(P*log(P))
  return(index)
}
```

```{r calculate shannon diversity}
plantDiversity <- plantDataEnvTotal %>%
  group_by(Environment) %>%
  summarize(H = shannon(Rel_abundance_env))
```

Plot Shannon diversity in each environment:

```{r plot diversity, echo=FALSE}
p2 <- ggplot(plantDiversity, aes(x = Environment, y = H, fill = Environment)) +
  geom_bar(stat = "identity", color = "grey40", alpha=.6) +
  theme_classic() +
  scale_fill_viridis_d() +
  labs(x = "", y = "Shannon's H")
ggsave("shannon.png", p2, dpi = 300)
p2
```

```{r}
```

```{r}
# Install and load required packages if not already installed
install.packages("kableExtra")
install.packages("webshot")
install.packages("htmlwidgets")
install.packages("scales")
install.packages("dplyr")

library(kableExtra)
library(webshot)
library(htmlwidgets)
library(scales)
library(dplyr)

overallRelativeAbundance <- plantData %>% 
  group_by(Common_name) %>%
  summarize(species_abundance = sum(Amount)) %>%
  mutate(total_abundance = 147, relative_abundance = species_abundance / total_abundance)

overallDiversity <- shannon(overallRelativeAbundance$relative_abundance)

overallRelativeAbundance <- overallRelativeAbundance %>%
  select(!c(species_abundance, total_abundance)) %>%
  mutate(relative_abundance = scales::percent(relative_abundance)) %>%
  rename(Species = Common_name, "Relative Abundance" = relative_abundance)

# Create the table
table_html <- overallRelativeAbundance %>%
  kbl(caption = "Overall relative abundance") %>%
  kable_classic(full_width = F, html_font = "Cambria")

# Save the table as an image
save_kable(table_html, "overallRelativeAbundance.png")


```

```{r}
```

```{r}
```

The total species richness in Oconee Forest Park (according to our limited study) is 47.

The overall Shannon diversity index is approximately 3.40

```{r}
library(ggsankey)
library(ggplot2)
library(dplyr)

Ethnobotany <- read.csv("native-plants-medicine.csv")
names(Ethnobotany) <- c('Plant', 'Usage', 'For')

# Step 1
df <- Ethnobotany %>%
  make_long(Plant, Usage, For)

# Chart 1
pl <- ggplot(df, aes(x = x,
                     next_x = next_x,
                     node = node,
                     next_node = next_node,
                     fill = factor(node),
                     label = node)) +
  geom_sankey(flow.alpha = 0.5,
              node.color = "black",
              show.legend = FALSE,
              width = 0.35
              ) +
  geom_sankey_label(size = 2.5, color = "black", fill = "white", position = "identity") +
  theme_bw() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank()) +
  scale_fill_viridis_d(option = "inferno") +
  theme_sankey(base_size = 12)
pl + theme(plot.margin = margin(0, 0, 0, 0, "cm"))  # Adjust the values as needed

ggsave(
  filename = "ethnobotany.png",
  plot = pl,
  width = 30,
  height = 30,
  units = "cm",
  dpi = 300
)
pl

```
